Evaluating infectious disease forecasts \\ with allocation scoring rules

### 2024-08-08 alloscore_manuscript/manuscript.Rnw
update diff.tex

Figure \ref{fig:bisection} illustrates this bisection method in the context of the examples of Section \ref {sec:methods.overview}, where it quickly finds close approximations to the analytic solutions $\tau^{\star} = 1-e^{-1}, 1-e^{-2}$ (for $K= 5,10$ respectively) which, in this case, are available (c.f. \eqref{eqn:exp-quantile}).

<<bisection, fig.height=4, fig.width = 8, fig.pos="H", fig.cap = "Illustrations of bisection method for the examples of Section \\ref{sec:methods.overview}. In each plot the thinner black lines are the quantile functions of the exponential forecasts for Location 1, with $\\sigma_1=1$ (lower), and Location 2, with $\\sigma_2=4$ (upper), while the thicker black line gives the sum of these quantiles, which must meet the resource constraints of $K = 5$ on the left and $K=10$ on the right. The red segments show the iterates $\\tau_{M,j}$ and their associated total resource demands as given by the function \\texttt{allocate} from the package \\texttt{alloscore} applied to these examples and using an initial search interval $[0,1]$. (The $\\tau_{M,j}$ are in this case just sums of terms $\\pm 2^{-j}$.) The horizontal dashed lines show the resulting allocations (c.f. Figure \\ref{fig:exp_alloc_example} ).">>=
expexamp <- tibble(
  target_names = LETTERS[1:2],
  dist = "exp",
  rate = c(1,.25),
) %>% add_pdqr_funs(types = c("p", "d", "q"))

@@

p10 <- p + geom_hline(aes(yintercept = 10)) +
  geom_hline(aes(yintercept = c(2,8)), linetype = 2) +
  geom_segment(data = iters10, aes(x=taus, xend = taus, y=0, yend = qtot), color = "red", alpha = .5) +
  scale_x_continuous(limits = c(.45,1), expand = c(0,0), 
                     breaks = c(.5,.6,.7,.8,0.865,.9,1),
                     labels = c("0.5","0.6",
                                "0.7","0.8", 
                                expression(italic(atop(" ", 1 - e^{-2} %~~% 0.865)), "0.9", "1"))) +
  scale_y_continuous(limits = c(0,12), breaks = c(0,2,5,8,10,15), expand = c(0,0), 
                     name = element_blank())

>>>>

Figure \ref{fig:bisection} illustrates this bisection method in the context of the examples of Section \ref{sec:methods.overview}, where it quickly finds close approximations to the analytic solutions $\tau^{\star} = 1-e^{-1}, 1-e^{-2}$ (for $K= 5,10$ respectively) which, in this case, are available (c.f. \eqref{eqn:exp-quantile}).

<<bisection, fig.height=4, fig.width = 8, fig.pos="H", fig.cap = "Illustrations of bisection method for the examples of Section \\ref{sec:methods.overview}. In each plot the thinner black curves are the quantile functions of the exponential forecasts for Location 1, with $\\sigma_1=1$ (lower), and Location 2, with $\\sigma_2=4$ (upper), while the thicker black curve gives the sum of these quantiles, which must meet the resource constraints of $K = 5$ on the left and $K=10$ on the right. The red segments show the iterates $\\tau_{M,j}$ and their associated total resource demands as given by the function \\texttt{allocate} from the package \\texttt{alloscore} applied to these examples and using an initial search interval $[0,1]$. (The $\\tau_{M,j}$ are in this case just sums of terms $\\pm 2^{-j}$.) The horizontal dashed lines show the resulting allocations (c.f. Figure \\ref{fig:exp_alloc_example} ).">>=
expexamp <- tibble(
  target_names = LETTERS[1:2],
  dist = "exp",
  rate = c(1,.25),
) %>% add_pdqr_funs(types = c("p", "d", "q"))

@@

p10 <- p + geom_hline(aes(yintercept = 10)) +
  geom_hline(aes(yintercept = c(2,8)), linetype = 2) +
  geom_segment(data = iters10, aes(x=taus, xend = taus, y=0, yend = qtot), color = "red", alpha = .5) +
  scale_x_continuous(limits = c(.45,1), expand = c(0,0), 
                     breaks = c(.5,.6,.7,.8,0.865,.9,1),
                     labels = c("0.5","0.6",
                                "0.7","0.8", 
                                expression(italic(atop(" ", 1 - e^{-2} %~~% 0.865)), "0.9", "1"))) +
  scale_y_continuous(limits = c(0,12), breaks = c(0,2,5,8,10,15), expand = c(0,0), 
                     name = "  ")

### 2024-08-08 alloscore_manuscript/manuscript.tex
update diff.tex

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Wed Aug  7 16:57:25 2024
\begin{table}[ht]
\centering
\begin{tabular}{lrrrr}
  \hline
Model & AS & MWIS & IAS centered at 15k & IAS uniform \\ 
  \hline
\texttt{CU-select} & 669 & 133 & 774 & 326 \\ 
  \texttt{per-capita} & 865 & - & 1029 & 366 \\ 
  \texttt{COVIDhub-ensemble} & 873 & 159 & 1067 & 438 \\ 
  \texttt{USC-SI\_kJalpha} & 995 & 91 & 1216 & 1097 \\ 
  \texttt{JHUAPL-Gecko} & 1034 & 164 & 1141 & 418 \\ 
  \texttt{MUNI-ARIMA} & 1084 & 169 & 1248 & 440 \\ 
  \texttt{COVIDhub-trained\_ensemble} & 1089 & 169 & 1271 & 823 \\ 
  \texttt{COVIDhub-baseline} & 1175 & 170 & 1317 & 535 \\ 
  \texttt{JHUAPL-Bucky} & 1358 & 102 & 1566 & 1214 \\ 
  \texttt{JHUAPL-SLPHospEns} & 1540 & 129 & 1604 & 1102 \\ 
  \texttt{UVA-Ensemble} & 2469 & 213 & 2635 & 2494 \\ 
   \hline
\end{tabular}
\caption{For one illustrative week, a comparison of allocation scores (AS), mean weighted interval scores (MWIS), and two varieties of Integrated Allocation Scores (IAS), (see Section \ref{sec:ias_examp}). All metrics are shown for 10 models that made forecasts of hospital admissions for 2022-01-03. Results are sorted by AS. For all metrics, lower scores indicate better accuracy.} 
\label{tab:multi-k-scores}
\end{table}

@@

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Wed Aug  7 16:57:30 2024
\begin{table}[ht]
\centering
\begin{tabular}{lrrr}
  \hline
model & MAS & taMWIS & taMWIS rank \\ 
  \hline
\texttt{COVIDhub-ensemble} & 389 & 70 & 2 \\ 
  \texttt{per-capita} & 464 & - & - \\ 
  \texttt{COVIDhub-trained\_ensemble} & 483 & 87 & 4 \\ 
  \texttt{CU-select} & 502 & 81 & 3 \\ 
  \texttt{JHUAPL-SLPHospEns} & 526 & 67 & 1 \\ 
  \texttt{COVIDhub-baseline} & 594 & 93 & 5 \\ 
  \texttt{JHUAPL-Bucky} & 643 & 112 & 7 \\ 
  \texttt{MUNI-ARIMA} & 707 & 116 & 8 \\ 
  \texttt{JHUAPL-Gecko} & 929 & 110 & 6 \\ 
  \texttt{USC-SI\_kJalpha} & 1473 & 155 & 9 \\ 
   \hline
\end{tabular}
\caption{Time averaged mean weighted interval scores (taMWIS) and mean allocation scores (MAS) by model across 13 weeks. These results show the average performance across time for the nine models that submitted forecasts for every week from 2021-11-29 through 2022-02-21, as well
  as for a per-capita allocation. MWIS is not defined for the \texttt{per-capita} allocation which does not use forecasts.} 
\label{tab:multi-week-performance-summary}
\end{table}

@@

Figure \ref{fig:bisection} illustrates this bisection method in the context of the examples of Section \ref {sec:methods.overview}, where it quickly finds close approximations to the analytic solutions $\tau^{\star} = 1-e^{-1}, 1-e^{-2}$ (for $K= 5,10$ respectively) which, in this case, are available (c.f. \eqref{eqn:exp-quantile}).

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[H]
\includegraphics[width=\maxwidth]{figure/bisection-1} \caption{Illustrations of bisection method for the examples of Section \ref{sec:methods.overview}. In each plot the thinner black lines are the quantile functions of the exponential forecasts for Location 1, with $\sigma_1=1$ (lower), and Location 2, with $\sigma_2=4$ (upper), while the thicker black line gives the sum of these quantiles, which must meet the resource constraints of $K = 5$ on the left and $K=10$ on the right. The red segments show the iterates $\tau_{M,j}$ and their associated total resource demands as given by the function \texttt{allocate} from the package \texttt{alloscore} applied to these examples and using an initial search interval $[0,1]$. (The $\tau_{M,j}$ are in this case just sums of terms $\pm 2^{-j}$.) The horizontal dashed lines show the resulting allocations (c.f. Figure \ref{fig:exp_alloc_example} ).}\label{fig:bisection}
\end{figure}

>>>>

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Wed Aug  7 23:13:42 2024
\begin{table}[ht]
\centering
\begin{tabular}{lrrrr}
  \hline
Model & AS & MWIS & IAS centered at 15k & IAS uniform \\ 
  \hline
\texttt{CU-select} & 669 & 133 & 774 & 326 \\ 
  \texttt{per-capita} & 865 & - & 1029 & 366 \\ 
  \texttt{COVIDhub-ensemble} & 873 & 159 & 1067 & 438 \\ 
  \texttt{USC-SI\_kJalpha} & 995 & 91 & 1216 & 1097 \\ 
  \texttt{JHUAPL-Gecko} & 1034 & 164 & 1141 & 418 \\ 
  \texttt{MUNI-ARIMA} & 1084 & 169 & 1248 & 440 \\ 
  \texttt{COVIDhub-trained\_ensemble} & 1089 & 169 & 1271 & 823 \\ 
  \texttt{COVIDhub-baseline} & 1175 & 170 & 1317 & 535 \\ 
  \texttt{JHUAPL-Bucky} & 1358 & 102 & 1566 & 1214 \\ 
  \texttt{JHUAPL-SLPHospEns} & 1540 & 129 & 1604 & 1102 \\ 
  \texttt{UVA-Ensemble} & 2469 & 213 & 2635 & 2494 \\ 
   \hline
\end{tabular}
\caption{For one illustrative week, a comparison of allocation scores (AS), mean weighted interval scores (MWIS), and two varieties of Integrated Allocation Scores (IAS), (see Section \ref{sec:ias_examp}). All metrics are shown for 10 models that made forecasts of hospital admissions for 2022-01-03. Results are sorted by AS. For all metrics, lower scores indicate better accuracy.} 
\label{tab:multi-k-scores}
\end{table}

@@

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Wed Aug  7 23:13:46 2024
\begin{table}[ht]
\centering
\begin{tabular}{lrrr}
  \hline
model & MAS & taMWIS & taMWIS rank \\ 
  \hline
\texttt{COVIDhub-ensemble} & 389 & 70 & 2 \\ 
  \texttt{per-capita} & 464 & - & - \\ 
  \texttt{COVIDhub-trained\_ensemble} & 483 & 87 & 4 \\ 
  \texttt{CU-select} & 502 & 81 & 3 \\ 
  \texttt{JHUAPL-SLPHospEns} & 526 & 67 & 1 \\ 
  \texttt{COVIDhub-baseline} & 594 & 93 & 5 \\ 
  \texttt{JHUAPL-Bucky} & 643 & 112 & 7 \\ 
  \texttt{MUNI-ARIMA} & 707 & 116 & 8 \\ 
  \texttt{JHUAPL-Gecko} & 929 & 110 & 6 \\ 
  \texttt{USC-SI\_kJalpha} & 1473 & 155 & 9 \\ 
   \hline
\end{tabular}
\caption{Time averaged mean weighted interval scores (taMWIS) and mean allocation scores (MAS) by model across 13 weeks. These results show the average performance across time for the nine models that submitted forecasts for every week from 2021-11-29 through 2022-02-21, as well
  as for a per-capita allocation. MWIS is not defined for the \texttt{per-capita} allocation which does not use forecasts.} 
\label{tab:multi-week-performance-summary}
\end{table}

@@

Figure \ref{fig:bisection} illustrates this bisection method in the context of the examples of Section \ref{sec:methods.overview}, where it quickly finds close approximations to the analytic solutions $\tau^{\star} = 1-e^{-1}, 1-e^{-2}$ (for $K= 5,10$ respectively) which, in this case, are available (c.f. \eqref{eqn:exp-quantile}).

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[H]
\includegraphics[width=\maxwidth]{figure/bisection-1} \caption{Illustrations of bisection method for the examples of Section \ref{sec:methods.overview}. In each plot the thinner black curves are the quantile functions of the exponential forecasts for Location 1, with $\sigma_1=1$ (lower), and Location 2, with $\sigma_2=4$ (upper), while the thicker black curve gives the sum of these quantiles, which must meet the resource constraints of $K = 5$ on the left and $K=10$ on the right. The red segments show the iterates $\tau_{M,j}$ and their associated total resource demands as given by the function \texttt{allocate} from the package \texttt{alloscore} applied to these examples and using an initial search interval $[0,1]$. (The $\tau_{M,j}$ are in this case just sums of terms $\pm 2^{-j}$.) The horizontal dashed lines show the resulting allocations (c.f. Figure \ref{fig:exp_alloc_example} ).}\label{fig:bisection}
\end{figure}
